EGGs are adhesive!\\ An algebraic account of equality saturation

### 2025-09-26 jlamp/jlamp.tex
fine

		\end{tikzpicture}
	\end{center}
	There are no morphisms in $\tg$ between the last two and from the last two to the first one, therefore none of them can be terminal.
\end{rem}

\begin{rem}
	$\tg$ is not an adhesive category. In particular it does not have pushouts along all monos. For instance, if we take the three term graphs of the previous remark, then have two arrows
	$(?_{\{h\}}, \id{\{v\}})\colon \Delta_{\tg}(\{v\})\to (l_a, !_{V_{\mathcal{G}_a}})$ and $(?_{\{h\}}, \id{\{v\}})\colon \Delta_{\tg}(\{v\})\to (l_b, !_{V_{\mathcal{G}_a}})$ which cannot be completed to a square. Indeed if $(q, !_{V_\mathcal{H}})\colon \mathcal{H}\to \mathcal{G}^\Sigma$ is another term graph with $(g_E, g_V)\colon (l_a, !_{V_{\mathcal{G}}})\to (q, !_{V_\mathcal{H}})$ and $(k_E, k_V)\colon (l_a, !_{V_{\mathcal{G}}})\to (q, !_{V_\mathcal{H}})$  such that 
	\[(g_E, g_V)\circ (?_{\{h\}}, \id{\{v\}}) = (k_E, k_V)\circ (?_{\{h\}}, \id{\{v\}})\]
	then $g_V=k_V$ and
	\[t_{\mathcal{H}}(g_E(h))=g^\star_V(t_{\mathcal{G}}(h))=g_V^\star(\delta_v)=k^\star_V(\delta_V)=k^\star_V(t_{\mathcal{G}}(h))=t_{\mathcal{H}}(k_E(h))\]
	so that we also have $g_E=k_E$, but then
	\[
	a=l_a(h)=q(g_E(h))=q(k_E(h))=l_b(h)=b\]
\end{rem}
}

@@

\noindent
\begin{minipage}[l]{.83\linewidth}
	\begin{restatable}{lem}{comp}\label{prop:eqhyp_complete}
		Consider a diagram $F\colon \D \to \EqHyp$ and let $(E_D, V_D, Q_D, s_D, t_D, q_D)$ be the image of an object $D$. Then the following facts hold
		\begin{enumerate}
			\item $F$ has a colimit, which is preserved by $K$;
			\item consider a cone $(L, \{l_D\}_{D\in \D})$ limiting  for $K \circ F$ and let $((E, V), \{(\pi^D_E, \pi^D_V)\}_{D\in \D})$ be one for $T\circ F$, then $F$ has a limit $(E, V, Q, s, t, q)$ and there is a mono $m\colon Q\mto L$ such that the diagram on the right commutes for every $D\in \D$.
		\end{enumerate}
	\end{restatable}
\end{minipage}\hfill
\begin{minipage}[r]{.15\linewidth}
\xymatrix{V \ar@{>>}[d]_{q} \ar[r]_{\pi^D_V}& V_D \ar@{>>}[dd]^{q_D}\\Q \ar@{>.>}[d]_{m}&\\ L \ar[r]^{l_d} & Q_D}
\end{minipage}

@@

\begin{defi}
	Let $\mathcal{G}$ be an hypergraph, a \emph{labelled hypergraph with equivalence} is a pair $(\mathcal{H}, (f,g))$ where $\mathcal{H}$ is an object of $\EqHyp$ and $(f,g)$ a morphism $T(\mathcal{H})\to \mathcal{G}$ of $\hyp$. A \emph{morphism of labelled hypergraphs with equivalence} between $(\mathcal{H}, (f,g))$ and $(\mathcal{H}', (h,k))$ is an arrow $(a,b, c)\colon \mathcal{H}\to \mathcal{H}'$ such that $h\circ a = f$ and $k\circ b = g$.
%
	We denote the resulting category by $\EqHyps$.
\end{defi}

\noindent 
\begin{minipage}[l]{.8\linewidth}
		\setlength{\parindent}{1.5em}
		\begin{rem}Let $\Sigma$ be a single typed monoidal signature (cfr~\Cref{def:alg}). Since $\mathcal{G}_{\Sigma}$ has a single node, then to label another hypergraph $\mathcal{H}$ it is enough to provide a function $l\colon E_{\mathcal{H}}\to O_{\Sigma}$ such that the diagrams aside commute.

			Notice that, if $\Sigma$ is algebraic, then such a function exists only if $\mathcal{H}$ is algebraic too. In that case, since the image of $\lg_{V_{\mathcal{H}}}$ is simply the singleton $\{1\}$, the second diagram commutes by default.
		\end{rem}
\end{minipage} \hfill 
\begin{minipage}[r]{.15\linewidth}
	\xymatrix@R=8pt{E_{\mathcal{H}} \ar[r]_{l} \ar[d]_{s_{\mathcal{H}}}& O_{\Sigma} \ar[d]^{\ari_\Sigma}\\ V^\star_{\mathcal{H}} \ar[r]^{\lgh_{V_{\mathcal{H}}}} & \mathbb{N}}
	\xymatrix@R=8pt{E_{\mathcal{H}} \ar[r]_{l} \ar[d]_{t_{\mathcal{H}}}& O_{\Sigma} \ar[d]^{\coar_\Sigma}\\ V^\star_{\mathcal{H}} \ar[r]^{\lgh_{V_{\mathcal{H}}}} & \mathbb{N}}
	\end{minipage}

Let $(\mathcal{H}, l)$ be an object of $\EqHyps$: since $T$ has a right adjoint $R$ by \Cref{prop:forghyp}, $(f,g)\colon T(\mathcal{H})\to \mathcal{G}^\Sigma$ corresponds to the arrow $(f, g, !_{Q_{\mathcal{H}}})\colon \mathcal{H}\to R(\mathcal{G})$. It is immediate to see that this correspondence extends to an equivalence with the slice category over $R(\mathcal{G})$.

\begin{prop}\label{prop:slice}
$\EqHyps$ is equivalent to $\EqHyp / R(\mathcal{G})$.
\end{prop}

@@

 \begin{prop}
 	Let $M_{\eq}\colon \Eqlmo\to \EqHyp$ and $S_{\eq}\colon \EqsHyp \to \EqHyp$ be the inclusion functors. Moreover, let $\Gamma_{R(\mathcal{G})}\colon 1\to \EqHyp$ be the constant functor in $R(\mathcal{G})$. Then
 	\begin{enumerate}
 		\item $\lmo_{\eq, \mathcal{G}}$ is equivalent to $\comma{M_\eq}{\Gamma_{R(\mathcal{G})}}$;
 		\item $\shyp_{\eq, \mathcal{G}}$ is equivalent to $\comma{S_\eq}{\Gamma_{R(\mathcal{G})}}$; 
 		\item $\pshyp_{\mathcal{G}}$ is equivalent to $\pshyp_{\eq}/R(\sp(\pr(\mathcal{G})))$. 
 	\end{enumerate}
 \end{prop}

@@

We can  now use \Cref{thm:slice-functors} to get the following.

@@

\begin{defi}
	Let $\mathcal{G} = (E, V, Q, s, t, q)$ be a hypergraph with equivalence and $(S, \pi_1, \pi_2)$ a kernel pair for $q^\star \circ s$.
	We will say that $\mathcal{G}$ is an \emph{e-hypergraph} if $q^\star \circ t \circ \pi_1 = q^\star \circ t \circ \pi_2$.

	We will denote by $\egg$ the full subcategory of $\EqHyp$ whose objects are e-hypergraphs, and by $E\colon \egg \to \EqHyp$ the associated inclusion functor.
\end{defi}

@@

A first result that we need is that limits in $\egg$ are computed as in $\EqHyp$. Full proofs are in \Cref{proof:elim}.

\begin{restatable}{lem}{elim}\label{lem:elim}
	$\egg$ has all limits and $E$ creates them.
\end{restatable}

@@

Finally, we can turn our attention to the labelled context.

\begin{defi} Let $\mathcal{G}$ be an hypergraph. We say that an object $(\mathcal{H}, (f,g))$ of $\EqHyps$ is a \emph{labelled e-hypergraph} if $\mathcal{H}$ is an e-hypergraph. We define the category $\eggl$ as the full subcategory of $\EqHyps$ given by labelled e-hypergraphs. We denote by $E_\mathcal{G}$ the corresponding inclusion functor .
\end{defi}

From the definitions, we get at once the follosing result.

\begin{prop}\label{prop:varie}
	Let $\mathcal{G}$ be an hypergraph and $\Gamma_{R(\mathcal{G})}$ the costant functor in $R(\mathcal{G})$. Then the following facts hold
	\begin{enumerate}
		\item $\eggl$ is equivalent to $\comma{E}\Gamma_{R(\mathcal{G})}$;
		\item the functor $W_\mathcal{G}\colon \eggl\to \egg$ forgetting the labeling creates all colimits, pullbacks and equalizers. 
	\end{enumerate}
\end{prop}

Let $\pbe_\mathcal{G}$ be the class of morphisms in $\eggl$ whose image in $\egg$ lies in $\pbe$. Notice that $\pbe_\mathcal{G}$ is also the class of arrows whose image through $E_\mathcal{G}$ is in $\pbc_\mathcal{G}$.
%
By \Cref{thm:slice-functors} we get the following result.

@@

\begin{rem}
When $\mathcal{G}$ is the hypergraph associated to an algebraic signature, the object of  $\lelm$ coincides with what in \cite{uovadesive} are called \emph{e-termgraphs with equivalence}. 
\end{rem}

\begin{lem}Let $\mathcal{G}$ be an hypergraph and $\Gamma_{R(\mathcal{G})}$ the costant functor in $R(\mathcal{G})$. 
	\begin{enumerate}
		\item $\lelm$ is equivalent to $\comma{E\circ M_{\mathsf{e}}}{\Gamma_{R(\mathcal{G})}}$
		\item $\leps$ is equivalent to $\comma{E\circ S_{\mathsf{e}}}{\Gamma_{R(\mathcal{G})}}$
	\end{enumerate}
\end{lem}

Define $\epbl_{\mathcal{G}}$ and $\epbs_{\mathcal{G}}$ as the classes of arrows whose underlying morphisms are in $\epbl$ or in $\epbs$. The following result now follows at once.

@@

                 \node [below=.5cm, align=flush center,text width=2cm] at (v1)
                {
                	Figure $(1)$.
                };
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a1) at (-2.3, -0.5){$a$};
		\node[style=none] (a1att) at (-2.3, -0.5){};
		\node[style=node](v1) at (-1.4, -0.5){};
		\node[style=node](w1) at (-0.9, -0.5){};
		\node[style=small box](a2) at (-2.3, 0.5){$a$};
		\node[style=none] (a2att) at (-2.3, 0.5){};
		\node[style=node](v2) at (-1.4, 0.5){};
		\node[style=node](w2) at (-0.9, 0.5){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};

		      \node [below=.5cm, align=flush center,text width=2cm] at (-1.15,-0.5)
		{
			Figure $(2)$.
		};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(a1att.center) to (v1);
		\draw(a2att.center) to (v2);
		\draw[out=0, in=-120] (w1) to (divfst.center);
		\draw[out=0, in=120](w2) to (divsnd.center);
		\draw(divout.center) to (w);
	\end{pgfonlayer}
	\begin{pgfonlayer}{eqlayer}
		\draw[dashed, rounded corners] (-1.7, 0.2) rectangle (-0.6, 0.8);
		\draw[dashed, rounded corners] (-1.7, -0.8) rectangle (-0.6, -0.2);
	\end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a) at (-1.8, 0){$a$};
		\node[style=none] (aatt) at (-1.5, 0){};
		\node[style=node](v) at (-0.9, 0){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};
		      \node [below=.5cm, align=flush center,text width=2cm] at (-0.9,-0.5)
		{
			Figure $(3)$.
		};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(aatt.center) to (v);
		\draw[out=-60, in=-180] (v) to (divfst.center);
		\draw[out=60, in=-180](v) to (divsnd.center);
		\draw(divout.center) to (w);
	\end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=node](w1) at (-0.9, -0.5){};
		\node[style=small box](a2) at (-2.3, 0){$a$};
		\node[style=none] (a2att) at (-2.3, 0){};
		\node[style=node](v) at (-1.4, 0){};
		\node[style=node](w2) at (-0.9, 0.5){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};

>>>>

		\end{tikzpicture}
	\end{center}
	There are no morphisms in $\tg$ between the last two and from the last two to the first one, therefore none of them can be terminal.
\end{rem}
In particular it does not have pushouts along all monos. For instance, if we take the three term graphs of the previous remark, then have two arrows
	$(?_{\{h\}}, \id{\{v\}})\colon \Delta_{\tg}(\{v\})\to (l_a, !_{V_{\mathcal{G}_a}})$ and $(?_{\{h\}}, \id{\{v\}})\colon \Delta_{\tg}(\{v\})\to (l_b, !_{V_{\mathcal{G}_a}})$ which cannot be completed to a square. Indeed if $(q, !_{V_\mathcal{H}})\colon \mathcal{H}\to \mathcal{G}^\Sigma$ is another term graph with $(g_E, g_V)\colon (l_a, !_{V_{\mathcal{G}}})\to (q, !_{V_\mathcal{H}})$ and $(k_E, k_V)\colon (l_a, !_{V_{\mathcal{G}}})\to (q, !_{V_\mathcal{H}})$  such that 
	\[(g_E, g_V)\circ (?_{\{h\}}, \id{\{v\}}) = (k_E, k_V)\circ (?_{\{h\}}, \id{\{v\}})\]
	then $g_V=k_V$ and
	\[t_{\mathcal{H}}(g_E(h))=g^\star_V(t_{\mathcal{G}}(h))=g_V^\star(\delta_v)=k^\star_V(\delta_V)=k^\star_V(t_{\mathcal{G}}(h))=t_{\mathcal{H}}(k_E(h))\]
	so that we also have $g_E=k_E$, but then
	\[
	a=l_a(h)=q(g_E(h))=q(k_E(h))=l_b(h)=b\]
\end{rem}
}

@@

\noindent
\begin{minipage}[l]{.83\linewidth}
	\begin{restatable}{lem}{comp}\label{prop:eqhyp_complete}
		Consider a diagram $F\colon \D \to \EqHyp$ and let $(E_D, V_D, Q_D, s_D, t_D, q_D)$ be the image of an object $D$. Then the following facts hold
		\begin{enumerate}
			\item $F$ has a colimit, which is preserved by $K$;
			\item consider a cone $(L, \{l_D\}_{D\in \D})$ limiting  for $K \circ F$ and let $((E, V), \{(\pi^D_E, \pi^D_V)\}_{D\in \D})$ be one for $T\circ F$, then $F$ has a limit $(E, V, Q, s, t, q)$ and there is a mono $m\colon Q\mto L$ such that the diagram on the right commutes for every $D\in \D$.
		\end{enumerate}
	\end{restatable}
\end{minipage}\hfill
\begin{minipage}[r]{.15\linewidth}
\xymatrix{V \ar@{>>}[d]_{q} \ar[r]_{\pi^D_V}& V_D \ar@{>>}[dd]^{q_D}\\Q \ar@{>.>}[d]_{m}&\\ L \ar[r]^{l_D} & Q_D}
\end{minipage}

@@

\begin{defi}
	Let $\mathcal{G}$ be an hypergraph with equivalence, the category $\EqHyps$ of \emph{labelled hypergraph with equivalence}  is $\EqHyp/\mathcal{G}$.
\end{defi}

\begin{rem} The equivalence relation on the  nodes of the labelling hypergraph put constraints on the possible relations on the labelled hypergraphs. Indeed, consider the following hypergraphs with equivalences $\mathcal{G}$ and $\mathcal{H}$.

\noindent
\begin{minipage}[l]{.6\linewidth}
	\setlength{\parindent}{1.5em}
	Notice that there is no morphism $\mathcal{H}\to \mathcal{G}$. Indeed $\id{T(\mathcal{G})}$ is the unique morphism $T(\mathcal{G})\to T(\mathcal{H})$ while there are only two morphisms $f, g\colon Q_{\mathcal{H}}\rightrightarrows Q_{\mathcal{G}}$: the first one is constant in $v$, while the second one is constant in $w$.
\end{minipage}\hfill \qquad 
\begin{minipage}[r]{.4\linewidth}
\begin{tikzpicture}[baseline=(v1)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a1) at (-.5, 1){};
		\node[style=none] (a1att) at (-.5, 1){};
		\node[style=node](v1) at (-.5, 0){};
		\node[style=small box](a2) at (.5, 1){};
		\node[style=none] (a2in) at (.22, 1){};
		\node[style=none] (a2out) at (.78, 1){};
		\node[style=node](v2) at (.5, 0){};
		\node[style=none] (l1) at (-.5, 1.5){$a$};
		\node[style=none](l2) at (.5, 1.5){$b$};
		\node[style=none] (l1) at (-1, 0){$v$};
		\node[style=none](l2) at (1, 0){$w$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(a1att.south) to (v1);
		\draw[style=none, out=180, in=180](v2) to (a2in.west);
		\draw[style=none, out=0, in=0](v2) to (a2out.east);
	\end{pgfonlayer}
	\begin{pgfonlayer}{eqlayer}
		\draw[dashed, rounded corners] (-.8, -0.3) rectangle (-.2, 0.3);
		\draw[dashed, rounded corners] (.2, -0.3) rectangle (.8, 0.3);
	\end{pgfonlayer}
\end{tikzpicture}\qquad  
\begin{tikzpicture}[baseline=(v1)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a1) at (-.5, 1){};
		\node[style=none] (a1att) at (-.5, 1){};
		\node[style=node](v1) at (-.5, 0){};
		\node[style=small box](a2) at (.5, 1){};
		\node[style=none] (a2in) at (.22, 1){};
		\node[style=none] (a2out) at (.78, 1){};
		\node[style=node](v2) at (.5, 0){};
		\node[style=none] (l1) at (-.5, 1.5){$a$};
		\node[style=none](l2) at (.5, 1.5){$b$};
		\node[style=none] (l1) at (-1, 0){$v$};
		\node[style=none](l2) at (1, 0){$w$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(a1att.south) to (v1);
		\draw[style=none, out=180, in=180](v2) to (a2in.west);
		\draw[style=none, out=0, in=0](v2) to (a2out.east);
	\end{pgfonlayer}
	\begin{pgfonlayer}{eqlayer}
		\draw[dashed, rounded corners] (-.8, -0.3) rectangle (.8, 0.3);
	\end{pgfonlayer}
\end{tikzpicture}
\end{minipage}

 For both cases we have:
\[ f(q_{\mathcal{H}}(w))=v \neq w =q_{\mathcal{G}}(w)=q_\mathcal{G}(\id{V_{\mathcal{G}}}(w)) \qquad  g(q_{\mathcal{H}}(v))=w \neq v =q_{\mathcal{G}}(v)=q_\mathcal{G}(\id{V_{\mathcal{G}}}(v))  \]

	Intuitively, labelling with $\mathcal{G}$ force us to keep all the sorts (i.e.~the labellings of the nodes) distinguished, while labelling with $\mathcal{H}$ allows us to merge different sorts.
\end{rem}

\noindent 
\begin{minipage}[l]{.8\linewidth}
	\setlength{\parindent}{1.5em}
\begin{exa}
	Let $\Sigma$ be a single typed monoidal signature (cfr~\Cref{def:alg}). Since $\mathcal{G}_{\Sigma}$ has a single node, then to label another hypergraph $\mathcal{H}$ it is enough to provide a function $l\colon E_{\mathcal{H}}\to O_{\Sigma}$ such that the diagrams aside commute.

	Notice that, if $\Sigma$ is algebraic, then such a function exists only if $\mathcal{H}$ is algebraic too. In that case, since the image of $\lg_{V_{\mathcal{H}}}$ is simply the singleton $\{1\}$, the second diagram commutes by default.
\end{exa}	
\end{minipage} \hfill 
\begin{minipage}[r]{.15\linewidth}
	\xymatrix@R=8pt{E_{\mathcal{H}} \ar[r]_{l} \ar[d]_{s_{\mathcal{H}}}& O_{\Sigma} \ar[d]^{\ari_\Sigma}\\ V^\star_{\mathcal{H}} \ar[r]^{\lgh_{V_{\mathcal{H}}}} & \mathbb{N}}
	\xymatrix@R=8pt{E_{\mathcal{H}} \ar[r]_{l} \ar[d]_{t_{\mathcal{H}}}& O_{\Sigma} \ar[d]^{\coar_\Sigma}\\ V^\star_{\mathcal{H}} \ar[r]^{\lgh_{V_{\mathcal{H}}}} & \mathbb{N}}
\end{minipage}

@@

 \begin{prop}\label{prop:eti}
 	Let $\mathcal{G}$ be an hypegraph with equivalence and consider the inclusion functors $M_{\eq}\colon \Eqlmo\to \EqHyp$, $S_{\eq}\colon \EqsHyp \to \EqHyp$ and $S_{p, \eq}\colon \pshyp_{\eq}\to \EqHyp$. Let $\Gamma_{\mathcal{G}}\colon 1\to \EqHyp$ be the constant functor in $\mathcal{G}$. Then
 	\begin{enumerate}
 		\item $\lmo_{\eq, \mathcal{G}}$ is equivalent to $\comma{M_\eq}{\Gamma_{\mathcal{G}}}$;
 		\item $\shyp_{\eq, \mathcal{G}}$ is equivalent to $\comma{S_\eq}{\Gamma_{\mathcal{G}}}$; 
 		\item $\pshyp_{\mathcal{G}}$ is equivalent to $\comma{S_{p, \eq}}{\Gamma_{\mathcal{G}}}$. 
 	\end{enumerate}
 \end{prop}

 In particular, the previous  proposition, together with Lemmas \ref{lem:equ}, \ref{lem:coequ}, \ref{colim} and \Cref{lim}, yields the following

 \begin{cor}\label{cor:eti}
 $\lmo_{\eq, \mathcal{G}}$ is closed in $\hyp_{\eq, \mathcal{G}}$ under equalizers, binary product and pullbacks. $\pshyp_{\eq, \mathcal{G}}$ is closed in $\hyp_{\eq, \mathcal{G}}$ under pullbacks. Both are closed in  $\hyp_{\eq, \mathcal{G}}$ under pushouts along arrows in $\pbc_{\mathcal{G}}$.
 \end{cor}

@@

We can  now use \Cref{thm:slice-functors} together with Lemmas \ref{lem:equ} and \ref{lem:coequ} to get the following.

@@

\begin{defi}
Let $\mathcal{G} = (E, V, Q, s, t, q)$ be a hypergraph with equivalence, two  hyperedges $e, e'\in E$ are \emph{$q$-joined} if $q^\star(t(e'))=q^\star(t(e))$.

We will say that $\mathcal{G}$ is an \emph{e-hypergraph} if every two  hyperedges $e,e'$ such that
$q^\star(s(e'))=q^\star(s(e))$ are $q$-joined.
We will denote by $\egg$ the full subcategory of $\EqHyp$ whose objects are e-hypergraphs, and by $E\colon \egg \to \EqHyp$ the associated inclusion functor. 
\end{defi}

We can reformulate the definition of e-hypergraphs in two other ways. See \Cref{proof:eh} for a proof

\begin{restatable}{prop}{eh}\label{prop:eh}
Consider an hypergraph with equivalence $\mathcal{G} = (E, V, Q, s, t, q)$ and let $(A, \pi^A_1, \pi^A_2)$, $(B, \pi^B_1, \pi^B_2)$ be kernel pairs for, respectively, $q^\star\circ s$ and $q^\star\circ t$. Then the following are equivalent:
\begin{enumerate}
	\item $\mathcal{G}$ is an e-hypergraph;
	\item $q^\star\circ t\circ \pi^A_1=q^\star\circ t\circ \pi^A_2$;
	\item there exists a unique arrow $k\colon A\to B$ such that 
	\[\pi^B_1\circ k = \pi^A_1 \qquad \pi^B_2\circ k = \pi^A_2 \]
	\end{enumerate}\end{restatable}

@@

As a next step, we we need to prove that connected limits in $\egg$ are computed as in $\EqHyp$. Full proofs are in \Cref{proof:elim}.

\begin{restatable}{lem}{elim}\label{lem:elimc}
Let $F\colon \D\to \hyp_{\eq}$ be a connected diagram and let $F(D)$ be $(E_D, V_D, Q_D, s_D, t_D, q_D)$. Suppose that $(\mathcal{L}, \{(a_D, b_D, c_D)\}_{D\in \D})$ is a limiting cone for $F$, then for every hyperedges $e,e'\in E_{\mathcal{L}}$, if $a_D(e)$ and $a_D(e')$ are $q_D$-joined for every $D\in \D$, then $e$ and $e'$ are $q_{\mathcal{L}}$-joined.
\end{restatable}

\begin{restatable}{cor}{ecim}\label{lem:elim}
	$\egg$ has all connected limits and $E$ creates them.
\end{restatable}

@@

Finally, we can turn our attention to the labelled context. In such an environment, we want the equivalence relation to identify the target of two hyperedges whenever their sources are related \emph{and} their labels are the same. This justifies the following definition.

\begin{defi} Let $\mathcal{G}$ be an hypergraph with equivalence. We say that an object $(a,b,c)\colon \mathcal{H}\to \mathcal{G}$ of $\EqHyps$ is a \emph{labelled e-hypergraph} if every two hyperedges $e,e'\in E_{\mathcal{H}}$ such that
	\[q^\star_{\mathcal{H}}(s(e))=q^\star_{\mathcal{H}}(s(e')) \qquad a(e) = a(e')\]
	are $q_\mathcal{H}$-joined.

 We define the category $\eggl$ as the full subcategory of $\EqHyps$ given by labelled e-hypergraphs. We denote by $E_\mathcal{G}$ the corresponding inclusion functor .
\end{defi}

\noindent
\begin{minipage}[l]{.75\linewidth}
	\begin{exa}
		Let $\mathcal{G}$ be the graph aside. Notice that $\id{\mathcal{G}}$ is a labelled e-hypergraph even if its underlying hypergraphs with equivalence $\mathcal{G}$ is not in $\egg$.
	\end{exa}
\end{minipage} \hfill \begin{minipage}[r]{.21\linewidth}
	\begin{tikzpicture}[baseline=(v1)]
		\begin{pgfonlayer}{nodelayer}
			\node[style=small box](a1) at (-1.4, 0){};
			\node[style=none] (a1att) at (-1.4, 0){};
			\node[style=node](v1) at (-.5, 0){};
			\node[style=small box](a2) at (1.4, 0){};
			\node[style=none] (a2in) at (1.4, 0){};
			\node[style=node](v2) at (.5, 0){};
			\node[style=none] (l1) at (-1.4, .5){$a$};
			\node[style=none](l2) at (1.4, .5){$b$};
			\node[style=none] (l1) at (-.5, -.5){$v$};
			\node[style=none](l2) at (.5, -.5){$w$};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw(a1att) to (v1);
			\draw[style=none](v2) to (a2in.south);
		\end{pgfonlayer}
		\begin{pgfonlayer}{eqlayer}
			\draw[dashed, rounded corners] (-.8, -0.3) rectangle (-.2, 0.3);
			\draw[dashed, rounded corners] (.2, -0.3) rectangle (.8, 0.3);
		\end{pgfonlayer}
	\end{tikzpicture}
\end{minipage}

We begin by providing an equivalent characterization to labelled e-hypergraphs (see \Cref{proof:fh} for details).

\begin{restatable}{lem}{fh}\label{lem:fh}
Let $(a,b,c)\colon \mathcal{H}\to \mathcal{G}$ be an object of $\hyp_{\eq, \mathcal{G}}$ and let $(A, \pi^A_1, \pi^A_2)$, $(B, \pi^B_1, \pi^B_2)$ be kernel pairs for, respectively, $(q^\star\circ s_{\mathcal{H}}, a),(q^\star\circ t_{\mathcal{H}}, a)\colon E_{\mathcal{H}}\to Q_{\mathcal{H}}\times E_{\mathcal{G}}$. Then the following are equivalent:
\begin{enumerate}
	\item $(a,b,c)$ is a labelled e-hypergraph;
	\item $q^\star\circ t_{\mathcal{H}}\circ \pi^A_1=q^\star\circ t_{\mathcal{H}}\circ \pi^A_2$;
	\item there exists a unique arrow $k\colon A\to B$ such that 
	\[\pi^B_1\circ k = \pi^A_1 \qquad \pi^B_2\circ k = \pi^A_2 \]
\end{enumerate}
\end{restatable}

We want now to show that pullbacks and pushout in $\eggl$ are computed as in $\hyp_{\eq, \mathcal{G}}$. See \Cref{proof:varie} for a proof.

\begin{restatable}{lem}{varie}\label{prop:varie}
	Let $\mathcal{G}$ be an hypergraph with equivalence, then $\eggl$ is closed in $\hyp_{\eq, \mathcal{G}}$ under connected limits and pushouts alongs arrows in $\pbe_\mathcal{G}$, where $\pbe_\mathcal{G}$ is the class of morphisms in $\eggl$ whose image through $E_\mathcal{G}$ is in $\pbc_\mathcal{G}$. 
\end{restatable}

%
By \Cref{thm:slice-functors} we get the following result.

@@

\begin{rem}
When $\mathcal{G}$ is the hypergraph associated to an algebraic signature, the object of  $\lelm$ coincides with what in \cite{uovadesive} are called \emph{e-term graphs with equivalence}. 
\end{rem}

Putting together \Cref{cor:eti} with \Cref{prop:varie} we get the following.

\begin{lem}Let $\mathcal{G}$ be an hypergraph with equivalence, then $\lelm$ and $\leps$ are closed in $\hyp_{\eq, \mathcal{G}}$ under pullbacks and pushouts along arrows in $\pbc_{\mathcal{G}}$.
\end{lem}

Define $\epbl_{\mathcal{G}}$ and $\epbs_{\mathcal{G}}$ as the classes of arrows whose underlying morphisms are in $\pbc_{\mathcal{G}}$ or in $\pbc_{\mathcal{G}}$. The following result now follows at once.

@@

                 \node [below=.5cm, align=flush center,text width=2cm] at (v1)
                {
                	Figure $(1)$.
                };
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
        \begin{pgfonlayer}{eqlayer}
        	\draw[dashed, rounded corners] (-1.2, -0.8) rectangle (-.6, -0.2);
        	\draw[dashed, rounded corners] (-1.2, 0.2) rectangle (-.6, 0.8);
        	\draw[dashed, rounded corners] (.6, -0.3) rectangle (1.2, 0.3);
        \end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a1) at (-2.3, -0.5){$a$};
		\node[style=none] (a1att) at (-2.3, -0.5){};
		\node[style=node](v1) at (-1.4, -0.5){};
		\node[style=node](w1) at (-0.9, -0.5){};
		\node[style=small box](a2) at (-2.3, 0.5){$a$};
		\node[style=none] (a2att) at (-2.3, 0.5){};
		\node[style=node](v2) at (-1.4, 0.5){};
		\node[style=node](w2) at (-0.9, 0.5){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};
